---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
d <- read.csv("data/activity.csv")

library(dplyr)
library(ggplot2)

d$date <- as.Date(d$date, "%Y-%m-%d")

head(d)
```


## What is mean total number of steps taken per day?

```{r}
d_sum <- d %>% dplyr::group_by(date) %>% 
    dplyr::summarise(steps_per_day = sum(steps))

hist(d_sum$steps_per_day)
```

```{r}
d_mean_med <- d %>% dplyr::group_by(date) %>% 
    dplyr::summarise(daily_mean = mean(steps, na.rm = TRUE),
                     daily_med = median(steps))

d_mean_med %>% knitr::kable(digits = 1)
```

```{r}
# Mean/Median total steps per day
summary(d_sum$steps_per_day)
```


## What is the average daily activity pattern?
```{r}
ave_int <- d %>% 
    group_by(interval) %>% 
    summarise(ave_steps = mean(steps, na.rm = TRUE)) %>% 
    ungroup()

ave_int %>% 
    ggplot(aes(interval, ave_steps)) +
        geom_line() +
        geom_text(data = subset(ave_int, ave_steps == max(ave_steps)), 
                  aes(label = paste("Max is", max(ave_int$ave_steps), "\n In interval", interval)))

```


## Imputing missing values
```{r}
# NAs = 2304
summary(d)
```
```{r}
# NAs are converted to 0, no daily medians where larger than 0 above
# Store in new drm_na variable
drm_na <- d
drm_na$steps[is.na(drm_na$steps)] <- 0L

drm_na_sum <- drm_na %>% group_by(date) %>% 
    summarise(daily_total = sum(steps))
```


The histogram distribution is more left skewed than when the NA values are removed.
```{r}
hist(drm_na_sum$daily_total)
```

```{r}
summary(drm_na_sum$daily_total)
```
The mean and Median are lower when the NA entries are 0.


## Are there differences in activity patterns between weekdays and weekends?

```{r}
plot_data <- drm_na %>% 
    mutate(day = weekdays(date),
           type_of_day = if_else(day %in% c("Sunday", "Saturday"), "Weekend", "Weekday")) %>% 
    group_by(interval, type_of_day) %>% 
    summarise(`Number of Steps` = mean(steps))



plot_data %>% ggplot(aes(interval, `Number of Steps`)) +
    geom_line(color = "blue") +
    facet_wrap(~type_of_day, ncol = 1)
```

